### 2017-08-31 Work_report/Presentation/GCN/gcn_presentation.tex
Adding template code for latex presentation

\documentclass{beamer}
\usetheme{Boadilla}

\title{My Presentation}
\subtitle{Using Beamer}
\author{Joe Bloggs}
\institute{University of ShareLaTeX}
\date{\today}

\begin{document}
% document goes here

\begin{frame}
\titlepage
\end{frame}

\section{Section 1}
\subsection{sub a}

\begin{frame}
\frametitle{Title}
Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua.
\end{frame}

\begin{frame}
\frametitle{Outline}
\tableofcontents
\end{frame}

\subsection{sub b}
\begin{frame}
\frametitle{List}
\begin{itemize}
\item Point A
\item Point B
\begin{itemize}
\item part 1
\item part 2
\end{itemize}
\item Point C
\item Point D
\end{itemize}
\end{frame}

\section{section 2}
\begin{frame}
  \frametitle{Enumerate}
  \begin{enumerate}[I]
  \item Point A
  \item Point B
    \begin{enumerate}[i]
    \item part 1
    \item part 2
    \end{enumerate}
  \item Point C
  \item Point D
  \end{enumerate}

\end{frame}

\subsection{sub A}
\begin{frame}
\frametitle{Using Columns}
\begin{columns}
\column{0.5\textwidth}
\begin{description}
\item[API] Application Programming Interface
\item[LAN] Local Area Network
\item[ASCII] American Standard Code for Information Interchange
\end{description}

\column{0.5\textwidth}
\begin{description}
\item[API] Application Programming Interface
\item[LAN] Local Area Network
\item[ASCII] American Standard Code for Information Interchange
\end{description}

\end{columns}
\end{frame}

\end{document}