### 2017-08-31 Work_report/Presentation/GCN/gcn_presentation.tex
I think presentation turned out decent enough
hoping that the presentation would go nice. But do good things always happen

\documentclass{beamer}
\usetheme{Boadilla}
% \usepackage[a4paper, tmargin=1in, bmargin=1in]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{pdflscape}
\usepackage{listings}
\usepackage{hyperref}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{amssymb}

\mode<presentation>{
    \AtBeginSection[]
    {
    \begin{frame}[allowframebreaks]{Outline}
    \tableofcontents[currentsection]
    \end{frame}
    }
}
% \AtBeginSubsection[
%   {\frame<beamer>{\frametitle{Outline}
%     \tableofcontents[currentsection,currentsubsection]}}%
% ]%
% {
%   \frame<beamer>{
%     \frametitle{Outline}
%     \tableofcontents[currentsection,currentsubsection]}
% }

\title{Graph Convolution Networks}
% \subtitle{Using Beamer}
\author{Arka Sadhu}
\institute{IIT Bombay}
\date{\today}

\begin{document}
% document goes here

\begin{frame}
\titlepage
\end{frame}

\begin{frame}
\frametitle{Outline}
\tableofcontents
\end{frame}

\section{Graph Convolutional Networks}

% \subsection{Why GCN}
\begin{frame}
  \frametitle{Introduction to Graph Convolutional Networks}
  \begin{itemize}
  \item <1-> CNN are extremely efficient architectures for image and audio classification tasks.
  \item <1-> But CNN donot directly generalize to irregular domains such as graph.
  \item <2-> Want to generalize CNN to Graphs.
  \item <3-> Non-trivial because the distances are non-euclidean.
  % \item <3-> Some applications would be spatio-temporal
  \end{itemize}
\end{frame}

% \subsection{How to extend convolution to graphs?}
\begin{frame}
  \frametitle{Extending Convolutional to Graphs}
  There are two main approaches
  \begin{itemize}
  \item <1->Spatial Approach :\\
    Generalization of CNN in the spatial domain itself.
    \begin{itemize}
    \item <2-> Learning Convolutional Neural Networks for Graphs [ICML 2016].\cite{DBLP:journals/corr/NiepertAK16}
    \end{itemize}
  \item <3-> Spectral Approach :\\
    Using the frequency characterization of CNN and using that to generalize to Graphical domain
    \begin{itemize}
    \item <4->Spectral Networks and Deep Locally Connected Networks on Graphs [Bruna et al. ICLR 2014]. % \cite{DBLP:journals/corr/BrunaZSL13} %
    \item <4->Convolutional Neural Networks on Graphs with Fast Localized Spectral Filtering [Defferrard et al. NIPS 2016] (will be the main focus) % \cite{DBLP:journals/corr/DefferrardBV16}
    \item <4->Semi-Supervised Classification with Graph Convolutional Networks [Kipf et al. ICLR 2017] % \cite{DBLP:journals/corr/KipfW16}
    \end{itemize}
  \end{itemize}
\end{frame}

\section{Problems with Spatial Approach}
\begin{frame}
  \frametitle{Limitations of Spatial Approach}
  \begin{itemize}
  \item Can't exactly define a neighborhood because the distances are not uniform.
  \item Ordering of nodes is problem specific.
  \end{itemize}
  Hence for the remainder we discuss the Spectral Approach
\end{frame}

\section{Spectral Approach}
% \subsection{Basics of Spectral Approach}
\begin{frame}
  \frametitle{A Basic Formulation}
  \begin{itemize}
  \item Convolution in spectral (Fourier) domain is point wise multiplication.
  \item Fourier Basis is defined as the eigen basis of the laplacian operator.
  \item Can use Laplacian of a graph.
  \end{itemize}
\end{frame}

% \subsection{Problem Formulation}
\begin{frame}
  \frametitle{Defining the Problem on Graphs}
  \begin{itemize}
  \item <1-> A feature description $x_i$ for every node $i$; summarized in a $N x D$ feature matrix $X$ ($N$ : number of nodes, $D$ : number of input features)
  \item <2-> Adjacency Matrix $A$.
  \item <3-> Node level output $Z$ (an $N x F$ feature matrix, where $F$ = number of output features per node).
  \item <4-> Every neural network can then be written as a non-linear function.
    $$H^{(l+1)} = f(H^{l}, A)$$
  \end{itemize}
\end{frame}

% \subsection{Graph Laplacian}
\begin{frame}
  \frametitle{Brief overview of Graph Laplacian}
  Let $T$ denote the diagonal matrix with ($v$,$v$)-th entry having value $d_v$ : degree of vertex $v$.
  % $$L(u,v) = $$
  Define L-matrix as
  \[
    L(u,v) =
    \begin{cases}
      d_v &\quad\text{if u = v}\\
      -1 &\quad\text{if u and v are adjacent}\\
      0 &\quad \text{otherwise}\\
    \end{cases}
  \]
  And the Laplacian of the graph as
  \[
    \mathcal{L}(u,v) =
    \begin{cases}
      1 &\quad\text{if u = v}\\
      -\frac{1}{\sqrt{d_ud_v}} &\quad\text{if u and v are adjacent}\\
      0 &\quad \text{otherwise}\\
    \end{cases}
  \]
\end{frame}

\begin{frame}
  \frametitle{Graph Laplacian (contd.)}
  $$\mathcal{L} = T^{-1/2}LT^{1/2}$$
  With the convention $T^{-1}(v,v) = 0$ for $d_v = 0$.\\
  When G is k-regular,
  $$\mathcal{L} = I - \frac{1}{k}A$$
  For a general graph
  $$\mathcal{L} = I - T^{-1/2}AT^{1/2}$$
\end{frame}
\section{Spectral Networks and Deep Locally Connected Networks on Graphs}
\begin{frame}
  \frametitle{Spectral Networks and Deep Locally Connected Networks on Graphs}
  \begin{itemize}
  \item Mentions the use of both spatial and spectral construction.
  \item For the spectral part uses a spline and has k control points for it.
    $$g_{\theta}(\Lambda) = B\theta$$
    Here $B$ is the cubic B-spline basis and $\theta$ is a vector of control points.
  \item The datasets used (created) are quite interesting. Subsampled MNIST and MNIST on sphere to show how spectral networks can be used on graphs.
  \end{itemize}
\end{frame}

\section{CNN on Graphs with Fast Localized Spectral Filtering}
\subsection{Learning fast localized Spectral filters}
\begin{frame}
  \frametitle{Graph Fourier Transform}
  \begin{itemize}
  \item Laplcian of the graph is real symmetric positive semidefinite, and thus can be written as
    $$L = U \Lambda U^{T}$$
  \item Here $U = [u_0 .... u_{n-1}]$ is the fourier basis and $\Lambda = diag([\lambda_0...\lambda_{n-1}])$ are ordered real non-negative eigen values.
  \item Graph Fourier Transform of a signal $x$ is $\hat{x} = U^{T}x$.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Spectral filtering of graph signals}
  \begin{itemize}
  \item Defining convolution on graphs
    $$ x *_{G} y = U((U^T x) \odot (U^T y))$$
  \item Filtering by $g_{\theta}$
    $$ y = g_{\theta}(L)x = g_{\theta}(U \Lambda U^{T})x = Ug_{\theta}(\Lambda) U^T x $$
  \item A non-parametric filter (all parameters free) would be defined as
    $$g_{\theta}(\Lambda) = diag(\theta)$$
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Polynomial Parametrization}
  \begin{itemize}
  \item   Problem with non-parametric filters is that not localized (we want something like k-neighborhood) and therefore their learning complexity becomes $O(n)$. This can be overcomed with use of a Polynomial filter
    $$g_{\theta}(\Lambda) = \sum_{k = 0}^{K-1}\theta_k\Lambda^k$$
  \item The advantage we gain here is that nodes which are at a distance greater than $K$ away from the node $i$, at which the filter is applied, are not affected. Hence we have gained localization.
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Recursive formulation for fast filtering}
  \begin{itemize}
  \item Still cost to filter is high $O(n^2)$ because of multiplication with $U$ matrix.
  \item Therefore use recurrence relation of chebyshev polynomial instead.
    $$g_{\theta}(\Lambda) = \sum_{k=0}^{K-1}\theta_k T_K(\tilde{\Lambda})$$
    Here $\tilde{\Lambda}$ is scaled between $[-1,1]$.
  \item This allows us to compute $\bar{x_k} = T_K{\tilde{L}}x$. And Therefore
    $$ y = g_{\theta}(L)x = [\bar{x_0}...\bar{x_{k-1}}]\theta$$
  \item The cost is now $O(K|E|)$
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Learning filters}
  \begin{itemize}
  \item Trivial to show that backprop calculation can be done efficiently.
  \end{itemize}
\end{frame}

\subsection{Coarsening and Pooling}
\begin{frame}
  \frametitle{Graph Coarsening and Pooling}
  \begin{itemize}
  \item Require efficient mechanism for pooling. Graph clustering as such is NP-hard and some approximations must be made.
  \item The paper uses Graclus algorithm for coarsening, and uses an intelligent way of rearranging the nodes [creating a balanced binary tree from the remaining singleton and fake nodes] so that the pooling now becomes equivalent to pooling a regular 1D signal.
  \end{itemize}
\end{frame}

% \subsection{Results on MNIST}
\begin{frame}
  \frametitle{MNIST results}
  \begin{itemize}
  \item Achieves close to classical CNN accuracy.
  \item Pictures to be added.
  \end{itemize}
\end{frame}

\section{Semi-Supervised Classification with Graph Convolutional Networks}
\begin{frame}
  \frametitle{Fast approximate convolutions on the graph}
  \begin{itemize}
  \item Layerwise propagation rule as
    $$H^{(l+1)} = \sigma(\tilde{D}^{-1/2}\tilde{A}\tilde{D}^{-1/2}H^{(l)}W^{(l)})$$
    Here $\tilde{A} = A + I_N$ the adjacency matrix with added self-loops. $\tilde{D}$ is the degree matrix.
  \item In the chebyshev approximation, limit to $K=1$ and therefore the layer-wise convolution operation is linear function of the laplacian.
  \item Experimentally shows 2-3 layered GCN can effectively learn standard graph problems. Specifically it does decently well in the unsupervised case, and significantly good in the semi-supervised setting.
  \end{itemize}
\end{frame}

\section{Limitations of Graph Convolutional Networks}
\begin{frame}
  \frametitle{Limitations}
  \begin{itemize}
  \item We would like to model a network which doesn't require a rectangular input size and therefore be able to accomodate superpixels. In theory this should be possible, but there are a few points that we need to keep in mind.
    \begin{itemize}
    \item Graph doesn't have orientation. There is no sense of up, down, left or right. The filters are rotationally invariant. This can be both advantageous as well as disadvantageous depending on the set-up of the problem. Spatial Transformer Networks learn the invariance to rotation as well as generic warping. But there is always the problem of '6' and '9' because they are equivalent in modulo rotation.
    \item The filters are not directly transferable to another graph (because of the graph laplacian).
    \end{itemize}
  \end{itemize}
\end{frame}

% \bibliography{ref.bib}
% \bibliographystyle{ieeetr}

\end{document}